---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figure/"
)
```

## gganatogram

Create anatogram images for different organisms. 
For now only human male is available.
This package uses the tissue coordinates from the figure in ArrayExpress Expression Atlas.
https://www.ebi.ac.uk/gxa/home

## Install

Install from github using devtools. 

```{r,eval=FALSE}
## install from Github
devtools::install_github("jespermaag/gganatogram")
```

## Usage

This package requires `ggplot2` and `ggpolypath`

```{r,eval=TRUE}
library(ggplot2)
library(ggpolypath)
library(gganatogram)
library(dplyr)
```

In order to use the function gganatogram, you need to have a data frame with
organ, colour, and value if you want to.

```{r, eval=TRUE}
organPlot <- data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
 type = c("circulation", "circulation",  "nervous system", "nervous system", "digestion", "digestion", "digestion"), 
 colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(10, 5, 1, 8, 2, 5, 5), 
 stringsAsFactors=F)

 head(organPlot)
 ```

Using the function gganatogram with the filling the oorgans based on colour.
```{r, eval=TRUE, organPlot, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="colour")
 ```

Of course, we can use the ggplot themes and functions to adjust the plots
```{r, eval=TRUE, organPlotvoid, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") + theme_void()
 ```

We can also skip plotting the outline of the graph
```{r, eval=TRUE, organPlotSubset, fig.width = 3, dpi = 100}
organPlot %>%
    dplyr::filter(type %in% c('circulation', 'nervous system')) %>%
gganatogram(outline=F, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") + theme_void()
 ```