---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figure/"
)
```

## gganatogram

Create anatogram images for different organisms.  <br/>
For now only human male is available.  <br/>
This package uses the tissue coordinates from the figure in ArrayExpress Expression Atlas.  <br/>
[https://www.ebi.ac.uk/gxa/home](https://www.ebi.ac.uk/gxa/home)  <br/>
[https://github.com/ebi-gene-expression-group/anatomogram](https://github.com/ebi-gene-expression-group/anatomogram)  <br/>

More plot examples can be found at [https://jespermaag.github.io/blog/2018/gganatogram/](https://jespermaag.github.io/blog/2018/gganatogram/)

## Install

Install from github using devtools. 

```{r,eval=FALSE}
## install from Github
devtools::install_github("jespermaag/gganatogram")
```

## Usage

This package requires `ggplot2` and `ggpolypath`

```{r,eval=TRUE}
library(ggplot2)
library(ggpolypath)
library(gganatogram)
library(dplyr)
```

In order to use the function gganatogram, you need to have a data frame with
organ, colour, and value if you want to.

```{r, eval=TRUE}
organPlot <- data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
 type = c("circulation", "circulation",  "nervous system", "nervous system", "digestion", "digestion", "digestion"), 
 colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(10, 5, 1, 8, 2, 5, 5), 
 stringsAsFactors=F)

 head(organPlot)
 ```

Using the function gganatogram with the filling the organs based on colour.
```{r, eval=TRUE, organPlot, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="colour")
 ```

Of course, we can use the ggplot themes and functions to adjust the plots
```{r, eval=TRUE, organPlotvoid, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") + 
theme_void()
```

We can also plot all tissues available using 
```{r, eval=TRUE, organPlotAll, fig.width = 3, dpi = 100}
gganatogram(data=hgMale_key, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") +theme_void()
 ```



We can also skip plotting the outline of the graph
```{r, eval=TRUE, organPlotSubset, fig.width = 3, dpi = 100}
organPlot %>%
    dplyr::filter(type %in% c('circulation', 'nervous system')) %>%
gganatogram(outline=F, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") + 
theme_void()
 ```


We can fill the tissues based on the values given to each organ
```{r, eval=TRUE, organPlotValue, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="value") + 
theme_void() +
scale_fill_gradient(low = "white", high = "red")
 ```

 We can also use facet_wrap to compare groups.  
 First create add two data frames together with different values and the conditions in the type column
```{r, eval=TRUE}
compareGroups <- rbind(data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
  colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(10, 5, 1, 8, 2, 5, 5), 
 type = rep('Normal', 7), 
 stringsAsFactors=F),
 data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
  colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(5, 5, 10, 8, 2, 5, 5), 
 type = rep('Cancer', 7), 
 stringsAsFactors=F))
```

```{r, eval=TRUE, Condition, fig.width = 6, dpi = 100}
gganatogram(data=compareGroups, fillOutline='#a6bddb', organism='human', sex='male', fill="value") + 
theme_void() +
facet_wrap(~type) +
scale_fill_gradient(low = "white", high = "red") 

 ```