---
output:
  md_document:
    variant: markdown_github
---



<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev='svg',
  fig.path = "figure/"
)
```

## gganatogram

```{r, include=FALSE}
if (requireNamespace("git2r")) {
  repo = git2r::repository(".")
  url = git2r::remote_url(repo, remote = "origin") 
  gh_user = basename(dirname(url))
} else {
  gh_user = "jespermaag"
}
```

[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/jespermaag/gganatogram?branch=master&svg=true)](https://ci.appveyor.com/project/`r gh_user`/gganatogram)
[![Travis build status](https://travis-ci.com/jespermaag/gganatogram.svg?branch=master)](https://travis-ci.com/`r gh_user`/gganatogram)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.1434233.svg)](https://doi.org/10.5281/zenodo.1434233)
  

Create anatogram images for different organisms.  <br/>
This package uses the tissue coordinates from the figure in Expression Atlas. [https://www.ebi.ac.uk/gxa/home](https://www.ebi.ac.uk/gxa/home)  <br/>
[https://github.com/ebi-gene-expression-group/anatomogram](https://github.com/ebi-gene-expression-group/anatomogram)  <br/>

## Citation
#### Maag JLV. gganatogram: An R package for modular visualisation of anatograms and tissues based on ggplot2 [version 1; referees: awaiting peer review]. F1000Research 2018, 7:1576 (doi: 10.12688/f1000research.16409.1)
https://f1000research.com/articles/7-1576/v1

```{r}
citation("gganatogram")
```

If you use gganatogram please cite Expression Atlas as well. <br/>
[Petryszak et al. 2015](https://academic.oup.com/nar/article/44/D1/D746/2502589) <br/>
Petryszak, Robert, Maria Keays, Y. Amy Tang, Nuno A. Fonseca, Elisabet Barrera, Tony Burdett, Anja Füllgrabe et al. "Expression Atlas update—an integrated database of gene and protein expression in humans, animals and plants." Nucleic acids research 44, no. D1 (2015): D746-D752.

More plot examples can be found at [https://jespermaag.github.io/blog/2018/gganatogram/](https://jespermaag.github.io/blog/2018/gganatogram/)

## Install

Install from github using devtools. 

```{r,eval=FALSE}
## install from Github
devtools::install_github("jespermaag/gganatogram")
```

## Usage

This package requires `ggplot2` and `ggpolypath` which loads when loading the package

```{r,eval=TRUE, message=FALSE}

library(gganatogram)
library(dplyr)
```
Available organisms 
```{r, eval=TRUE, AllSpeciesPlot, fig.width = 10, dpi = 300}
library(gridExtra)
hgMale <- gganatogram(data=hgMale_key, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") + theme_void()
hgFemale <- gganatogram(data=hgFemale_key, fillOutline='#a6bddb', organism='human', sex='female', fill="colour") + theme_void()
mmMale <- gganatogram(data=mmMale_key, fillOutline='#a6bddb', organism='mouse', sex='male', fill="colour") + theme_void()
mmFemale <- gganatogram(data=mmFemale_key, outline = T, fillOutline='#a6bddb', organism='mouse', sex='female', fill="colour")  +theme_void()  

grid.arrange(hgMale, hgFemale, mmMale, mmFemale, ncol=4)
 ```


In order to use the function gganatogram, you need to have a data frame with
organ, colour, and value if you want to.

```{r, eval=TRUE}
organPlot <- data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
 type = c("circulation", "circulation",  "nervous system", "nervous system", "digestion", "digestion", "digestion"), 
 colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(10, 5, 1, 8, 2, 5, 5), 
 stringsAsFactors=F)

 head(organPlot)
 ```

Using the function gganatogram with the filling the organs based on colour.
```{r, eval=TRUE, organPlot, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="colour")
 ```

Of course, we can use the ggplot themes and functions to adjust the plots
```{r, eval=TRUE, organPlotvoid, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") + 
theme_void()
```

We can also plot all tissues available using hgMale_key
```{r, eval=TRUE, organPlotAll, fig.width = 3, dpi = 100}
hgMale_key$organ
gganatogram(data=hgMale_key, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") +theme_void()
 ```



We can also skip plotting the outline of the graph
```{r, eval=TRUE, organPlotSubset, fig.width = 3, dpi = 100}
organPlot %>%
    dplyr::filter(type %in% c('circulation', 'nervous system')) %>%
gganatogram(outline=F, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") + 
theme_void()
 ```


We can fill the tissues based on the values given to each organ
```{r, eval=TRUE, organPlotValue, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="value") + 
theme_void() +
scale_fill_gradient(low = "white", high = "red")
 ```

 We can also use facet_wrap to compare groups.  
 First create add two data frames together with different values and the conditions in the type column
```{r, eval=TRUE}
compareGroups <- rbind(data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
  colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(10, 5, 1, 8, 2, 5, 5), 
 type = rep('Normal', 7), 
 stringsAsFactors=F),
 data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
  colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(5, 5, 10, 8, 2, 5, 5), 
 type = rep('Cancer', 7), 
 stringsAsFactors=F))
```

```{r, eval=TRUE, Condition, fig.width = 6, dpi = 100}
gganatogram(data=compareGroups, fillOutline='#a6bddb', organism='human', sex='male', fill="value") + 
theme_void() +
facet_wrap(~type) +
scale_fill_gradient(low = "white", high = "red") 

 ```

You can also split the tissues into types while retaining the outline 
```{r, eval=TRUE, maleOrgans, fig.width = 10, fig.height = 8,dpi = 200}
gganatogram(data=hgMale_key, outline = T, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") +
facet_wrap(~type, ncol=4) +
theme_void()

 ```

## Added female option
All female tissues
```{r, eval=TRUE, female, fig.width = 3,  dpi = 200}
hgFemale_key$organ
gganatogram(data=hgFemale_key, outline = T, fillOutline='#a6bddb', organism='human', sex='female', fill="colour")  +theme_void()
 ```

You can also split the tissues into types while retaining the outline 
```{r, eval=TRUE, femaleOrgans, fig.width = 10, fig.height = 8, dpi = 200}
gganatogram(data=hgFemale_key, outline = T, fillOutline='#a6bddb', organism='human', sex='female', fill="colour") +
facet_wrap(~type, ncol=4) +
theme_void()

 ```

To display the female reproductive system with outline.
```{r, eval=TRUE, femaleRep, fig.width = 6,  dpi = 200}
hgFemale_key %>%
    dplyr::filter(type=='reproductive') %>%
    gganatogram( outline = T, fillOutline='#a6bddb', organism='human', sex='female', fill="colour")  +
    theme_void()  +
    coord_cartesian(xlim = c(30, 75), ylim = c(-110, -80))

 ```

# Added  mouse

## Male
```{r, eval=TRUE, maleMouse, figwidth = 3, dpi = 50}
mmMale_key$organ
gganatogram(data=mmMale_key, outline = T, fillOutline='#a6bddb', organism='mouse', sex='male', fill="colour")  +theme_void()  

```

```{r, eval=TRUE, maleMouseOrgan, figwidth = 10, figheight = 8, dpi = 200}

gganatogram(data=mmMale_key, outline = T, fillOutline='#a6bddb', organism='mouse', sex='male', fill="colour")  +theme_void()+facet_wrap(~type, ncol=4)
```

## Female
```{r, eval=TRUE, femaleMouse, figwidth = 3, dpi = 50}
mmFemale_key$organ
gganatogram(data=mmFemale_key, outline = T, fillOutline='#a6bddb', organism='mouse', sex='female', fill="colour")  +theme_void()  

```

```{r, eval=TRUE, femaleMouseOrgan, figwidth = 10, figheight = 8, dpi = 200}

gganatogram(data=mmFemale_key, outline = T, fillOutline='#a6bddb', organism='mouse', sex='female', fill="colour")  +theme_void()+facet_wrap(~type, ncol=4)
```


## Other organisms i.e. tier 2 organisms
Expression atlas contains other organisms than human and mice, however, these are not as well anotated.  <br/>
All the expression atlas anatograms can be found here https://ebi-gene-expression-group.github.io/anatomogram/   <br/>
Unfortunately, I won't be able to add other organs to these since I'm neither an anatomist nor artist.    <br/>
If anyone would like to add more organs, I would love for you to contribute.  <br/>
  <br/>
To create these plots, I have added two other objects other_key and other_list.  <br/>
These are lists within lists, and to plot all the organs from an organisms use other_key[["organism"]] as data, and "organism" as organism.  <br/>
Also, the organ names are so far a mix of UBERON and plant ids. 

```{r}
length(other_key)
names(other_key)

```

To plot bos_taurus use the following command.
Unfortunately, I have not managed to add the correct names yet.

```{r,  eval=TRUE, bosTaurus}
other_key[["bos_taurus"]]

gganatogram(data=other_key[["bos_taurus"]], outline = T, fillOutline='white', organism="bos_taurus", sex='female', fill="colour")  +
        theme_void() +
        ggtitle("bos_taurus") + 
        theme(plot.title = element_text(hjust=0.5)) + 
        coord_fixed()
```

Here is a way to loop through all the other organisms and plot their organs.

```{r, eval=TRUE, othersFirst12, figwidth = 12, figheight = 16, dpi = 200}
library(gridExtra)
plotList <- list()
for (organism in names(other_key)) {
    plotList[[organism]] <- gganatogram(data=other_key[[organism]], outline = T, fillOutline='white', organism=organism, sex='female', fill="colour")  +
                theme_void() +
                ggtitle(organism) + 
                theme(plot.title = element_text(hjust=0.5, size=9)) + 
                coord_fixed()
}

do.call(grid.arrange,  c(plotList[1:4], ncol=2))

do.call(grid.arrange,  c(plotList[5:8], ncol=2))

do.call(grid.arrange,  c(plotList[9:12], ncol=2))

do.call(grid.arrange,  c(plotList[13:16], ncol=2))

do.call(grid.arrange,  c(plotList[17:20], ncol=2))

do.call(grid.arrange,  c(plotList[21:24], ncol=2))


```