---
output:
  md_document:
    variant: markdown_github
---



<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev='svg',
  fig.path = "figure/"
)
```

## gganatogram

```{r, include=FALSE}
if (requireNamespace("git2r")) {
  repo = git2r::repository(".")
  url = git2r::remote_url(repo, remote = "origin") 
  gh_user = basename(dirname(url))
} else {
  gh_user = "jespermaag"
}

```

[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/`r gh_user`/gganatogram?branch=master&svg=true)](https://ci.appveyor.com/project/`r gh_user`/gganatogram)
[![Travis build status](https://travis-ci.com/`r gh_user`/gganatogram.svg?branch=master)](https://travis-ci.com/`r gh_user`/gganatogram)  
  

Create anatogram images for different organisms.  <br/>
For now only human male is available.  <br/>
This package uses the tissue coordinates from the figure in Expression Atlas. [https://www.ebi.ac.uk/gxa/home](https://www.ebi.ac.uk/gxa/home)  <br/>
[https://github.com/ebi-gene-expression-group/anatomogram](https://github.com/ebi-gene-expression-group/anatomogram)  <br/>

If you use gganatogram please cite Expression Atlas as well <br/>
[Petryszak et al. 2015](https://academic.oup.com/nar/article/44/D1/D746/2502589)
Petryszak, Robert, Maria Keays, Y. Amy Tang, Nuno A. Fonseca, Elisabet Barrera, Tony Burdett, Anja Füllgrabe et al. "Expression Atlas update—an integrated database of gene and protein expression in humans, animals and plants." Nucleic acids research 44, no. D1 (2015): D746-D752.

More plot examples can be found at [https://jespermaag.github.io/blog/2018/gganatogram/](https://jespermaag.github.io/blog/2018/gganatogram/)

## Install

Install from github using devtools. 

```{r,eval=FALSE}
## install from Github
devtools::install_github("jespermaag/gganatogram")
```

## Usage

This package requires `ggplot2` and `ggpolypath` which loads when loading the package

```{r,eval=TRUE, message=FALSE}

library(gganatogram)
library(dplyr)
```

In order to use the function gganatogram, you need to have a data frame with
organ, colour, and value if you want to.

```{r, eval=TRUE}
organPlot <- data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
 type = c("circulation", "circulation",  "nervous system", "nervous system", "digestion", "digestion", "digestion"), 
 colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(10, 5, 1, 8, 2, 5, 5), 
 stringsAsFactors=F)

 head(organPlot)
 ```

Using the function gganatogram with the filling the organs based on colour.
```{r, eval=TRUE, organPlot, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="colour")
 ```

Of course, we can use the ggplot themes and functions to adjust the plots
```{r, eval=TRUE, organPlotvoid, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") + 
theme_void()
```

We can also plot all tissues available using hgMale_key
```{r, eval=TRUE, organPlotAll, fig.width = 3, dpi = 100}
hgMale_key$organ
gganatogram(data=hgMale_key, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") +theme_void()
 ```



We can also skip plotting the outline of the graph
```{r, eval=TRUE, organPlotSubset, fig.width = 3, dpi = 100}
organPlot %>%
    dplyr::filter(type %in% c('circulation', 'nervous system')) %>%
gganatogram(outline=F, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") + 
theme_void()
 ```


We can fill the tissues based on the values given to each organ
```{r, eval=TRUE, organPlotValue, fig.width = 3, dpi = 100}
gganatogram(data=organPlot, fillOutline='#a6bddb', organism='human', sex='male', fill="value") + 
theme_void() +
scale_fill_gradient(low = "white", high = "red")
 ```

 We can also use facet_wrap to compare groups.  
 First create add two data frames together with different values and the conditions in the type column
```{r, eval=TRUE}
compareGroups <- rbind(data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
  colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(10, 5, 1, 8, 2, 5, 5), 
 type = rep('Normal', 7), 
 stringsAsFactors=F),
 data.frame(organ = c("heart", "leukocyte", "nerve", "brain", "liver", "stomach", "colon"), 
  colour = c("red", "red", "purple", "purple", "orange", "orange", "orange"), 
 value = c(5, 5, 10, 8, 2, 5, 5), 
 type = rep('Cancer', 7), 
 stringsAsFactors=F))
```

```{r, eval=TRUE, Condition, fig.width = 6, dpi = 100}
gganatogram(data=compareGroups, fillOutline='#a6bddb', organism='human', sex='male', fill="value") + 
theme_void() +
facet_wrap(~type) +
scale_fill_gradient(low = "white", high = "red") 

 ```

You can also split the tissues into types while retaining the outline 
```{r, eval=TRUE, maleOrgans, fig.width = 10, fig.height = 8,dpi = 200}
gganatogram(data=hgMale_key, outline = T, fillOutline='#a6bddb', organism='human', sex='male', fill="colour") +
facet_wrap(~type, ncol=4) +
theme_void()

 ```

## Added female option
All female tissues
```{r, eval=TRUE, female, fig.width = 3,  dpi = 200}
hgFemale_key$organ
gganatogram(data=hgFemale_key, outline = T, fillOutline='#a6bddb', organism='human', sex='female', fill="colour")  +theme_void()
 ```

You can also split the tissues into types while retaining the outline 
```{r, eval=TRUE, femaleOrgans, fig.width = 10, fig.height = 8, dpi = 200}
gganatogram(data=hgFemale_key, outline = T, fillOutline='#a6bddb', organism='human', sex='female', fill="colour") +
facet_wrap(~type, ncol=4) +
theme_void()

 ```

To display the female repdoductive system with outline.
```{r, eval=TRUE, femaleRep, fig.width = 3,  dpi = 200}
hgFemale_key %>%
    dplyr::filter(type=='reproductive') %>%
    gganatogram( outline = T, fillOutline='#a6bddb', organism='human', sex='female', fill="colour")  +
    theme_void()  +
    coord_cartesian(xlim = c(30, 75), ylim = c(-110, -80))

 ```
